
# Try MOSAIKS {#sec-intro-demo}

```{r}
#| echo: false
#| results: "asis"

library(tibble)

source("../../_common.R")
status("testing_en")
```

## Overview

This chapter includes a link to a python notebook demonstration that replicates key results from the original MOSAIKS publication @rolf_etal_2021. While MOSAIKS has great potential to improve access to satellite-based prediction in data-sparse environments, the original paper focused on demonstrating performance in the United States where high-quality training data was readily available.

The US served as an ideal testing ground for several reasons:

- Extensive ground truth data available across multiple variables
- Reliable spatial referencing of data
- Diverse landscapes and built environments
- Ability to benchmark against existing methods
- Systematic validation of predictions

This validation in a data-rich environment was crucial for establishing MOSAIKS as a reliable tool before deploying it in contexts where ground truth data is scarce or unreliable. While the publication establishes the validity of the method, the following demonstration notebook provides a practical guide for users to showcase how to use MOSAIKS to generate novel predictions for a variety of variables. 

## Demonstration code

### Workflow

Below is a link to a Jupyter notebook intended to demonstrate practical use of MOSAIKS with real data. In fact, this notebook uses the original input data and features from @rolf_etal_2021. The code demonstrates:

1. Loading pre-computed MOSAIKS features and labels
2. Merging the features and labels
3. Training a ridge regression model
4. Evaluating predictions
5. Visualizing results

### Label data

The demo showcases several variables which we call "labels." These labels are the target variables that MOSAIKS aims to predict. The variables include:

```{r}
#| echo: false
#| results: "asis"

labels <- readr::read_csv("../tables/rolf_variables.csv", show_col_types = FALSE)

# Build the tabset using knit_child with inline text
panels <- lapply(seq_len(nrow(labels)), function(i) {
    row <- labels[i, ]
    knitr::knit_child(text = c(
        glue::glue("## {row$variable}"),
        "",
        glue::glue("![**Left maps:** 80,000 observations used for training and validation, aggregated up to 20 km × 20 km cells for display. **Right maps:** concatenated validation set estimates from 5-fold cross-validation for the same 80,000 grid cells (observations are never used to generate their own prediction), identically aggregated for display. **Scatters:** Validation set estimates (vertical axis) vs. ground truth (horizontal axis); each point is a ~1 km × 1 km grid cell. Black line is at 45°. **Data:** {row$description}.]({row$image}){{#{row$fig_id}}}"),
        ""
    ), envir = environment(), quiet = TRUE)
})

cat("::: {.panel-tabset}\n\n")
cat(unlist(panels), sep = "\n")
cat("\n:::")
```

A user simply needs to select which variable they would like to predict, and no other changes need to be made to the code. All data has been preprocessed and the code will download the necessary files from Zenodo. Users can also modify the code to investigate other modeling, evaluation, and visualization approaches. Any changes can be saved to a personal copy of the notebook in Google Drive. 

### Constraints

To stay within the Google Colab free tier (which imposes limits on memory usage), we subset the data. We take a 50% random sample of both features (*K*=4,000 instead of 8,192) and observations (*N*=50,000 instead of 100,000) compared to the original paper. Despite using this reduced dataset, the demo still achieves strong predictive performance, highlighting MOSAIKS's efficiency.

## Run the code!

::: {.callout-note}
# Click the badge to run the demonstration! 

&darr;&darr;&darr;&darr;&darr;&darr;&darr;&darr;&darr;&darr;&darr;&darr;&darr;&darr;&darr;&darr;   
[![Open In Colab](https://colab.research.google.com/assets/colab-badge.svg)](https://colab.research.google.com/github/center-for-effective-global-action-cega/MOSAIKS-Training-Manual/blob/main/notebooks/demo/01-Rolf_et_al_2021_demo.ipynb)  
<!-- &uarr;&uarr;&uarr;&uarr;&uarr;&uarr;&uarr;&uarr;&uarr;&uarr;&uarr;&uarr;&uarr;&uarr;&uarr;&uarr;     -->
Remember to click `File` -> `Save a copy in Drive` to save any changes you make.  
---
Or to view a static version of the code on GitHub, click the badge below.  
[![](https://img.shields.io/badge/View%20on%20GitHub-181717?logo=github&logoColor=white&labelColor=grey&color=black&link=--)](https://github.com/center-for-effective-global-action-cega/MOSAIKS-Training-Manual/blob/main/notebooks/demo/01-Rolf_et_al_2021_demo.ipynb)
:::

For instructions and tips on using Google Colab, please refer to @sec-intro-compute. 

## Don't want to run code?

Consider watching this demonstration instead!

::: {#fig-demo}

{{< video https://www.youtube.com/embed/ux2c-ot3Ce0?si=Oalkj-4MsJN5hNCh >}}

An overview of MOSAIKS and a live demonstration of generating novel predictions using the system. Video recorded by CIGAR Generalized Planetary Remote Sensing - 2020 Convention session. Presented by Esther Rolf and Tamma Carleton.
:::

## What's next? 

After establishing MOSAIKS's capabilities in the US context, the MOSAIKS development team have successfully demonstrated the system in many additional settings. This includes on the global scale, or in settings with few or low quality data. In the coming chapters, we will explore some of these applications, showing how MOSAIKS can help address data gaps in regions where traditional data collection is challenging or costly.

::: {.callout-note}
# Looking forward

In the next section we will take a closer look at the label data that can be used with MOSAIKS. 
:::
